Towards Verifying the Bitcoin-S Library

### 2019-06-29 paper/paper.tex
more work

\begin{abstract}
  We try to verify properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in subset of Scala called
  the \emph{Pure Scala Fragment}. We first try to verify the property
  that regular transactions do not create new money. It turns out that
  there is too much code involved that lies outside of the supported
  fragment to make this feasible. However, in the process we uncover
  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
  interesting) property: that adding zero satoshis to a given amount
  of satoshis yields the given amount of satoshis. Here as well a
  significant part of the relevant code lies outside of the supported
  fragment. However, after a series of equivalent transformations we
  arrive at code that we successfully verify.

@@

\paragraph{The Stainless Verifier.} Stainless is the successor of the
Leon verifier
\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
and is developed at EPF Lausanne
\cite{Stainless:github}. It is intended to be
used by programmers without training in formal verification and thus
allows to write specifications in Scala and focusses on counterexample
finding in addition to proving correctness.

The example in Figure~\ref{fig:factorial} from the Stainless documentation
\cite{Stainless:documentation} demonstrates this. Notice how a
precondition is specified using the function \emph{require} and a
postcondition using \emph{ensuring}.

\begin{figure}
\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}
	\caption{Factorial program with specification}
	\label{fig:factorial}
\end{figure}
Our program happens not to satisfy the specification. An overflow in
the 32-bit Int leads to a negative result for the input 17, as
Stainless reports in Figure~\ref{fig:failed}. Changing the type from
Int to BigInt will result in a successful verification.

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{output1.png}
	\caption{Output of Stainless verification for calculating factorial of Int number}
	\label{fig:failed}
\end{figure}

@@

What will turn out to be more relevant are the following Scala
features which Stainless does not support, such as: (concrete) class
definitions, inheritance by objects, abstract type members, variance
annotations and private inner classes.

In addition, Stainless has its own library of some core data types and
functions which need to be mapped correctly to functions inside of the
SMT solver that Stainless ultimately relies on. Those data types in
general do not have all the methods of the Scala data types. For
example, the BigInt type in Scala has a methods for bitwise operations
while the BigInt type in Stainless does not.

\paragraph{Outline and Properties to Verify.} In the next section we
try to verify the property that a regular (non-coinbase) transaction
cannot generate new coins. We call it the \emph{no-inflation
  property}. Trying to verify it, we uncover and fix a bug in the
bitcoin-s library. We then find that there is too much code involved
that lies outside of the supported fragment to currently make this
verification feasible. Instead, we turn to a simpler property to
verify. The simplest possible property we can think of is the fact
that adding zero satoshis to a given amount of satoshis yields the
given amount of satoshis. We call it the \emph{addition-with-zero
  property} and we try to verify it in Section~3. Here as well we see
that a significant part of the code lies outside of the supported
fragment. However, after a series of equivalent transformations we
arrive at code that we successfully verify.

@@

A crucial function for the verification of the no-inflation property
is the \texttt{checkTransaction} function, shown in
Figure~\ref{fig:checktrans}. To better understand it, we first see how
to create a transaction.

\paragraph{Creating a Transaction}

% The bitcoin-s library has a bitcoin
% transaction builder class with the following constructor:

% \begin{lstlisting}[style=scala]
% def apply(
%       destinations: Seq[TransactionOutput],
%       utxos: BitcoinTxBuilder.UTXOMap,
%       feeRate: FeeUnit,
%       changeSPK: ScriptPubKey,
%       network: BitcoinNetwork): Future[BitcoinTxBuilder] = {...}
% \end{lstlisting}

% The purpose of the return type Future is not clear to us here, since
% the constructor always returns a completed future. This might be for
% future purposes.

To create a transaction, we first need some coins -- an unspent
transaction output. We could load an actual unspent transaction output
from the bitcoin network, but we create one manually in order to see
this process. So we first create an (invalid) transaction with one
output in Figure~\ref{fig:prevtx}.

@@

We first create a keypair, then a lock script with its public key,
then the amount of satoshis, then a transaction output (utxo) for that
amount and locked with that script. Finally we create the actual
transaction with that output and no inputs. Of course, that is not a
valid transaction, because it creates coins out of nothing. In
particular, checkTransaction(prevTx) is false.

Now that we have a transaction output, we create a new transaction to
spend it.

First, we need some out points.  They point to outputs of previous
transactions.  We use the index zero, because the previous transaction
has only one output that becomes the first index zero.  If there were
two previous outputs, the second output would become the index 1 and
so on.
\begin{figure}
\begin{lstlisting}[style=scala]
  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)

@@

  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
    destinations = destinations, // where to send the money
    utxos = utxos,               // unspent transaction outputs
    feeRate = feeRate,           // fee rate per byte
    changeSPK = creditingSPK,    // where to send the change
    network = networkParams      // bitcoin network information
  )

@@

  val tx: Transaction = Await.result(signedTxF, 1 second)
\end{lstlisting}
  \caption{Creating a transaction}
  \label{fig:tx}
\end{figure}

This utxos are the inputs of our transaction.
Second, we need destinations to spend the bitcoins to.
For the sake of convenience we create only one.
We spend 5000 satoshis to the newly created random public key.
Finally, we define the fee rate in satoshis per one byte transaction
size as well as some bitcoin network parameters.  The bitcoin network
parameters are not important, so we use some static values normally
used when testing.

Now lets build the transaction with those data.
Line one to seven creates a transaction builder which is then signed
on line ten.  We can now use our transaction object on line twelve.
For example, after calling \emph{hex} on it, we can send the returned
string to the bitcoin network.

\paragraph{Validating a Transaction}

Bitcoin-S offers a function called \emph{checkTransaction} located in the ScriptInterpreter object.

  \caption{The \texttt{checkTransaction} function in the \texttt{ScriptInterpreter} object}
  \label{fig:checktrans}
\end{figure}

We can pass a transaction and it returns a Boolean indicating whether
the transaction is valid or not.  So for example when we pass the
transaction we built before the returned value would be true, because
it's a valid transaction.  It might not be accepted by the bitcoin
network but for a transaction on its own it's valid.  We can not check
context with it, because we can only pass one transaction.

There are several checks in checkTransaction.
For example, it checks if there is either no input or no output.
In this case we get false.

The relevant part for the bug we found:
\begin{lstlisting}[style=scala]
  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
\end{lstlisting}

It gathers all transaction ids referenced by the out points.  When we
call \emph{distinct} on the returned list, we get a list with
duplicate removed.  If the size of the new list is the same as the
size of the old, we know that there was no duplicate transaction id,
because, as said, distinct removes the duplicates.

\paragraph{Trying Stainless on the entire project.}

\begin{itemize}
\item result sbt: no output
\item result jar: 
\end{itemize}

In order to verify a project, Stainless must be integrated into it.
We can integrate it in an sbt project adding the Stainless Plugin 
and the required resolver to pugins.sbt. 
Another option to use Stainless is to import its libraries in a program code 
and verify a program from command line using the pre-packaged Stainless JAR file 
or using the Stainless script built from source.
Trying to integrate Stainless in Bitcoin-S caused a lot of troubles,
mainly because of version conflicts.  For more details see chapter
\ref{chap:appendix_arb}.

After integrating the Stainless plugin in the Bitcoin-S sbt project, 
there were many errors because of the different sbt versions.  
Some errors are described in the section 4.2.
It takes too much time to fix them all so it should be easier to extract
the classes needed for the \texttt{checkTransaction} function.

\paragraph{Putting aside the No-Inflation Property.} Naively trying
Stainless on the entire bitcoin-s codebase results in either no output
(with sbt) or many errors (with jar). So we extract the relevant code
to only verify that. However, the extracted code has more than 1500
lines and liberally uses Scala features outside of the supported
fragment.  We tried to transform the code into the supported fragment,
but realize that a better approach is to first verify a simpler
property with less code involved and then turn back to the
no-inflation property with more experience. So we turn to the
addition-with-zero property in the next section. \todo{explain}

\paragraph{Fixing a Bug in Bitcoin-S}

We can see that there is a bug in the checkTransaction function from
before, recognized and fixed through this work.

Here is the relevant code of checkTransaction again:
\begin{lstlisting}[style=scala]
  val prevOutputTxIds = transaction.inputs.map(_.previousOutput.txId)
  val noDuplicateInputs = prevOutputTxIds.distinct.size == prevOutputTxIds.size
\end{lstlisting}

What happens if we have two TransactionOutPoints (previousOutputs) with a different index but referencing the same Transaction ID (txId)?

According to the Bitcoin protocol this is possible.  A transaction can
have multiple outputs that should be referenceable by the next
transaction.  So this is clearly a bug.

What should not be possible is a transaction referencing the same
output twice.  This bug occurred in Bitcoin Core known as
CVE-2018–17144 which was patched on September 18,
2018. \cite{cve201817144}

Here, Bitcoin-S did a bit too much and marked all transaction as
invalid, if they referenced the same transaction twice.  The fix is,
to check on TransactionOutPoint instead of TransactionOutPoint.txId,
because TransactionOutPoint contains the txId as well as the output
index it references.  So in pseudo code, we check on the tuple (tx,
index) instead of (tx).  The fixed code:
\begin{lstlisting}[style=scala]
  val prevOutputs = transaction.inputs.map(_.previousOutput)
  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
\end{lstlisting}

Since TransactionOutPoint is a case class and Scala has a built in ==
for case classes there is no need to implement TransactionOutPoint.==.

This was fixed in
\href{https://github.com/bitcoin-s/bitcoin-s/pull/435}{pull request
  number 435} on GitHub at April 23, 2019, through this work along
with a unit test to prevent this bug from appearing again in the
future.

@@

\label{chap:verify_add}

In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.
We look at the verification of the addition of Satoshis with zero Satoshis.
This operation should result in the same amount of Satoshis.
Let's call it the \nameref{property_2}.

Using Stainless, we see the successful verification of this property.
But the process of the verification with the tool requires many
changes in the code, so that Stainless can accept it.  We look at all
needed modifications in chapter \ref{chap:verify_add}.

After realizing that it would consume too much time to rewrite the
Bitcoin-S code and even the extracted part with checkTransaction, the
smallest unit in Bitcoin-S-Core that is worthwhile to verify was
extracted.  This could be the addition of two CurrencyUnits.  To make
it even easier, the addition of CurrencyUnits with zero.
CurrencyUnits is an abstract class in Bitcoin-S, representing
currencies like Satoshis.

\paragraph{Extracting the relevant Code}

The relevant code is in two files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.

reference: code/addition/src/main/scala/addition/original

From those files we removed all code that is not needed for our
verification. For example, we removed all number types except for
Int64 (so Int32, UInt64, etc.) because we do not use them. We also
removed the superclasses Factory and NetworkElement of CurrencyUnit
and Number, respectively because the inherited members are not used by
the relevant code. Also, we removed all binary operations on Number that are not used.

  - removed extending Factory, NetworkElement and BasicArithmetic. This includes some hex/byte conversion eg fromHex, hex, bytes, fromBytes, ... Just interfaces never referenced
    \href{https://bitcoin-s.org/docs/core/core-intro\#the-basics}{description in section \#the-basics (cannot add link with hashtag)}
    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/protocol/NetworkElement.scala}{NetworkElement class}
    \href{https://github.com/bitcoin-s/bitcoin-s/blob/master/core/src/main/scala/org/bitcoins/core/util/Factory.scala}{Factory class}

  - removed Bitcoins class

  - removed subtraction and multiplication, binary operations <<, >>, etc, comparision operator <=, >=, etc but not == and !=

  - removed toBigDecimal

  - removed object CurrencyUnits containing some variables to transform satoshis to btc (not used)

  The code we use for the following sections is in reduced folder.

Here we can see the extracted code needed for the addition of CurrencyUnits:
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}

@@

When we run Stainless on this code (without any properties to prove),
it throws the following errors: \todo{describe errors: no support for
  abstract types, unsupported arguments for the BigInt constructor,
  unsupported inheritance for objects.}

@@

\todo{kai todo}

@@

As specified, our verification must only support addition with zero.
So we restrict the parameter to be zero in the precondition.
\begin{lstlisting}[style=scala]
  require(c.satoshis == Satoshis.zero)
\end{lstlisting}

We ensure the result is the same value as \emph{this} in the postcondition.
\begin{lstlisting}[style=scala]
  ensuring(res => res.satoshis == this.satoshis)
\end{lstlisting}

@@

That's all we need to verify our addition.

@@

The verified code.
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/number/NumberType.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/modified/currency/CurrencyUnits.scala}

@@

\appendix
\section{Project Setup}

\paragraph{Using the Stainless-sbt-plugin vs Using the Stanless-JAR}

@@

Why JAR:
sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)

@@

\subsection{Integration into Bitcoin-S}

During this work, Stainless updated the sbt plugin to support sbt
1.2.8 from 0.13.17 and Scala 2.12.8 from 2.11.12.  So this section
might be out of date now.

To use the latest version of the sbt tool you have to build it
locally.  You can run \texttt{sbt universal:stage} in the cloned
Stainless git repository.  This generates
\emph{frontends/scalac/target/universal/stage/bin/stainless-scalac}.

Bitcoin-S-Core uses sbt 1.2.8 and Scala 2.12.8, while Stainless sbt
plugin is on sbt 0.13.17 and Scala 2.11.12.

Sbt introduced new features in the 1.x release used by Bitcoin-S.
Most of them can be written the sbt 0.13.17 way.

The bigger problem is, due to the different Scala and sbt versions,
the following error after trying to go in a sbt shell:
\begin{verbatim}
  [warn] There may be incompatibilities among your library dependencies; run 'evicted'
         to see detailed eviction warnings.
  [error] java.lang.NoClassDefFoundError: sbt/SourcePosition
  ...
  Project loading failed: (r)etry, (q)uit, (l)ast, or (i)gnore?
\end{verbatim}

Downgrading Bitcoin-S sbt version to 0.13.17 fixes the error but then
it can not load some libraries only compiled for newer versions.  So
this would take too much time to fix and changes the Bitcoin-S code
inadvertently.

The next approach is to use the stainless cli instead of sbt.  Running
stainless on all source files does not work, because dependencies are
missing.  The parameter \emph{-classpath} can resolve it but the value
of this parameter must be the paths to all the dependencies separated
by a ':'.  Finally, \emph{core} depends on \emph{secp256k1jni},
another package of Bitcoin-S written in Java.  So this needs to be in
the source files to.

The final command looks like this in \emph{core} folder of Bitcoin-S:
\begin{lstlisting}[language=bash]
  $ stainless
    -classpath ".:$(find ~/.ivy2/ -type f -name *.jar | tr '\n' ':')"
    $(find . -type f -name *.scala | tr '\n' ' ')
    $(find ../secp256k1jni -type f -name *.java | tr '\n' ' ')
\end{lstlisting}

\emph{.ivy2} is the dependency cache of sbt.  The \emph{tr} replaces
the first char with the second so a newline with either ':' or ' '.

With this command, Stainless throws the next error:
\begin{verbatim}
  [Internal] Error: object scala.reflect.macros.internal.macroImpl in compiler mirror
             not found.. Trace:
  [Internal] - scala.reflect.internal.MissingRequirementError$.signal
             (MissingRequirementError.scala:17)
  ...
  [Internal] object scala.reflect.macros.internal.macroImpl in compiler mirror not found.
  [Internal] Please inform the authors of Inox about this message
\end{verbatim}

So we can not know how many errors will face us.  Let's go another
way, because the errors may take too much time and it might lead to a
next error.  We extract the code needed to verify a transaction mainly
the class Transaction and ScriptInterpreter with many other classes
they're depending on.

After this extraction Stainless was successfully integrated with both
sbt and JAR.

Running \texttt{sbt compile} in the project with Stainless ended
without error.  But it also ended with no output.  So we are not able
to change the code so Stainless would accept it since we do not know
what to change.

So the sbt plugin does not always complain where the JAR file did.
The open
\href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
  GitHub} might describe exactly this error.

\section{NumberType.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/number/NumberType.scala}

\section{CurrencyUnits.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/addition/reduced/currency/CurrencyUnits.scala}

\section{Code Transformations}

@@

After:
\begin{lstlisting}[style=scala]
sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = underlying.toBigInt

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}
\end{lstlisting}

@@

>>>>

\begin{abstract}
  We try to verify properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in a subset of Scala
  called \emph{Pure Scala}. We first try to verify the property that
  regular transactions do not create new money. It turns out that
  there is too much code involved that lies outside of the supported
  fragment to make this feasible. However, in the process we uncover
  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
  interesting) property: that adding zero satoshis to a given amount
  of satoshis yields the given amount of satoshis. Here as well a
  significant part of the relevant code lies outside of the supported
  fragment. However, after a series of equivalent transformations we
  arrive at code that we successfully verify.

@@

\paragraph{The Stainless Verifier.} Stainless is the successor of the
Leon verifier
\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
and is developed at EPF Lausanne \cite{Stainless:github}. It is
intended to be used by programmers without training in formal
verification. To facilitate that, it accepts specifications written in
the programming language itself (Scala). Also, it focusses on
counterexample finding in addition to proving
correctness. Counterexamples are useful to programmers while
correctness proofs are not -- correctness is obvious or does not hold,
and often both at the same time.

The example in Figure~\ref{fig:factorial} adapted from the Stainless
documentation \cite{Stainless:documentation} shows how the verifier is
used. Notice how a precondition is specified using \emph{require} and
a postcondition using \emph{ensuring}.

\begin{figure}
\begin{lstlisting}[style=scala]
  def factorial(n: Int): Int = {
      require(n >= 0)
      if (n == 0) {
        1
      } else {
        n * factorial(n - 1)
      }
  } ensuring(res => res >= 0)
\end{lstlisting}
	\caption{Factorial program with specification}
	\label{fig:factorial}
\end{figure}
Our program does not satisfy the specification. An overflow in the
32-bit Int leads to a negative result for the input 17, as Stainless
reports in Figure~\ref{fig:failed}. Changing the type from Int to
BigInt will result in a successful verification.

\begin{figure}
	\centering
		\includegraphics[width=\textwidth]{output1.png}
	\caption{Stainless output for the factorial program}
	\label{fig:failed}
\end{figure}

@@

What will turn out to be more relevant are the Scala features which
Stainless does not support, such as: (concrete) class definitions,
inheritance by objects, abstract type members, and inner classes in
case objects.

Also, Stainless has its own library of some core data types and
functions which need to be mapped correctly to functions inside of the
SMT solver that Stainless ultimately relies on. Those data types in
general do not have all the methods of the Scala data types. For
example, the BigInt type in Scala has a methods for bitwise operations
while the BigInt type in Stainless does not.

\paragraph{Outline and Properties to Verify.} In the next section we
try to verify the property that a regular (non-coinbase) transaction
can not generate new coins. We call it the \emph{no-inflation
  property}. Trying to verify it, we uncover and fix a bug in the
bitcoin-s library. We then find that there is too much code involved
that lies outside of the supported fragment to currently make this
verification feasible. So we turn to a simpler property to verify. The
simplest possible property we can think of is the fact that adding
zero satoshis to a given amount of satoshis yields the given amount of
satoshis. We call it the \emph{addition-with-zero property} and we try
to verify it in Section~3. Here as well we see that a significant part
of the code lies outside of the supported fragment. However, after a
series of equivalent transformations we arrive at code that we
successfully verify.

@@

  \caption{The \texttt{checkTransaction} function}
  \label{fig:checktrans}
\end{figure}

A crucial function for the verification of the no-inflation property
is the function which validates transactions: the
\texttt{checkTransaction} function in the ScriptInterpreter object,
shown in Figure~\ref{fig:checktrans}. It takes a transaction and if
some basic checks succeed returns true, otherwise false. For example,
one of those checks is that both the list of inputs and list of
outputs need to be non-empty.

To better understand the validation of a transaction in bitcoin-s, it
is useful to review how transactions are represented and created.

\paragraph{Creating a Transaction.} The code in this subsection is
adapted from the bitcoin-s documentation. To create a transaction, we
first need some coins -- an unspent transaction output. We could load
an actual unspent transaction output from the bitcoin network, but we
create one manually in order to see this process. So we first create
an (invalid) transaction with one output in Figure~\ref{fig:prevtx}.

@@

We first create a keypair, then a lock script with its public key,
then the amount of satoshis, then a transaction output (utxo) for that
amount and locked with that script. Finally we create the actual
transaction with that output and no inputs. Of course, that is not a
valid transaction, because it creates coins out of nothing. In
particular, checkTransaction(prevTx) returns false, simply because the
list of inputs is empty.

\begin{figure}
\begin{lstlisting}[style=scala]
  val outPoint = TransactionOutPoint(prevTx.txId, UInt32.zero)

@@

  val txBuilderF: Future[BitcoinTxBuilder] = BitcoinTxBuilder(
    destinations = destinations, 
    utxos = utxos,               
    feeRate = feeRate,           
    changeSPK = creditingSPK,    // where to send the change
    network = networkParams      
  )

@@

  val tx: Transaction = Await.result(txF, 1 second)
\end{lstlisting}
  \caption{Creating a transaction}
  \label{fig:tx}
\end{figure}

Now that we have a transaction output, we create a transaction to
spend it.

First, we need a so-called \emph{out point}, which points to an output
of a previous transaction. The second parameter is the index of the
referenced output. Here we use zero because the previous transaction
has only one output (with index zero).

Second, we need some information on how to spend the utxo from the
previous transaction, in particular, how to sign the transaction to
allow this. That is the information needed to create a transaction
input, so essentially a transaction input.

Third, we assemble the list of transaction inputs, in our case just one.

Fourth, we set the amount of satoshis that we want to spend. The Int64
class emulates C data types in Scala, and we will look at it more in
the next section.

Fifth, we create a lock script to send the coins to. To be supremely
secure, we use a public key derived from a private key we immediately
forget.

Sixth, we create our list of transaction outputs.

Seventh and eighth, we define the fee rate and some bitcoin network parameters.

Nineth, we create a transaction builder with those data.

Tenth, we tell the transaction builder to start signing the transaction.

Finally, we get the actual signed transaction. We could serialize it
and send it to the Bitcoin network. We can also pass it to the
checkTransaction function, which will return true.

\paragraph{A Bug in the checkTransaction function.} Note lines 16 and
17 of the checkTransaction function in Figure~\ref{fig:checktrans}.

The value prevOutputTxIds gathers all transaction IDs referenced by
the inputs of the current transaction.  When we call \emph{distinct}
on the returned list, we get the list with duplicates removed.  If the
size of the new list is the same as the size of the old, we know that
there is no duplicate transaction ID.

However, this code prevents a transaction with two inputs spending two
different outputs of a previous transaction. The check is too strict:
checkTransaction returns false for valid transactions.

The fix is simple: we perform the duplicate check on the
TransactionOutPoints instead of on their transaction IDs.  A
TransactionOutPoint contains the txId as well as the output index it
references.

Specifically, we replace lines 16-17 as follows:
\begin{lstlisting}[style=scala, firstnumber=16]
  val prevOutputs = transaction.inputs.map(_.previousOutput)
  val noDuplicateInputs = prevOutputs.distinct.size == prevOutputs.size
\end{lstlisting}

Note that TransactionOutPoint is a case class and thus has a built in
== method.

We submitted the fix together with a unit test to prevent this bug
from appearing again in the future in a pull request
\cite{BitcoinS:pull435}, which has already been merged.

\paragraph{An Attempt at Verification.} Naively trying Stainless on
the entire bitcoin-s codebase results in many errors -- as was to be
expected. We tried to extract only the relevant code and to verify
that. However, even the extracted code has more than 1500 lines and
liberally uses Scala features outside of the supported fragment.  We
tried to transform the code into the supported fragment, but quickly
realized that that a better approach is to first verify a simpler
property with less code involved and later come back to the
no-inflation property with more experience. So we now turn to the
addition-with-zero property.

@@

In Bitcoin-S there is a class \texttt{Satoshis} representing an amount of bitcoins.

@@

\paragraph{Extracting the relevant Code} The relevant code is in two
files: \texttt{CurrencyUnits.scala} and \texttt{NumberType.scala}.

From those files we removed all code that is not needed for our
verification. For example, we removed all number types except for
Int64 (so Int32, UInt64, etc.) because we do not use them. We also
removed the superclasses Factory and NetworkElement of CurrencyUnit
and Number, respectively, because the inherited members are not used by
the relevant code. Also, we removed all binary operations on Number
that are not used, like subtraction and multiplication.

@@

When we run Stainless on this code (without any properties to prove),
it throws the following errors: describe errors: no support for
  abstract types, unsupported arguments for the BigInt constructor,
  unsupported inheritance for objects.}

\todo{...}

@@

@@

@@

@@

\clearpage
\appendix

\section{Project Setup}

\paragraph{Using the Stainless-sbt-plugin vs Using the Stainless-JAR}

@@

Why JAR:
sbt-plugin problems (v 0.1 uses older versions (v0.2 not), no output to console see github \#484)
otherwise use sbt-plugin because of library management and replicability (everyone uses the same Stainless version configured in sbt)
no sbt because

-- version conflicts
-- no output maybe \href{https://github.com/epfl-lara/stainless/issues/484}{issue 484 on
  GitHub}

@@

\section{Extracted Code}
\label{sec:extracted}

\subsection{NumberType.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/number/NumberType.scala}

\subsection{CurrencyUnits.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/reduced/currency/CurrencyUnits.scala}

\section{Code Transformations} 
\label{sec:transform}

@@

@@

\section{Verified Code}
\label{sec:verified}

\subsection{NumberType.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/number/NumberType.scala}

\subsection{CurrencyUnits.scala}
\lstinputlisting[style=scala]{../code/addition/src/main/scala/verified/currency/CurrencyUnits.scala}
